Full Analysis Script

### 2024-04-18 code/analysis-code/fullanalysis-v1.qmd
Project Part 5
Finished ML analysis and redid tables in final manuscript.

rm(list=ls()) #clear the environment
```

@@

@@

whereas BW is birthweight in grams, MA is maternal ancestry (Euro-descent, Afro-descent), and $\epsilon$ is the error term. Results are displayed in Table 2.

```{r}
# Model using recipes and workflows
# Define recipe for model
ML_recipe1 <- recipe(Weightgrams ~ ModifiedColor, data = ML_linear) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable

@@

# Save regression output into table
saveRDS(table2, file = here::here("results", "tables", "table2_final.rds"))
```

@@

whereas BW is birthweight in grams, MA is maternal ancestry (Euro-descent, Afro-descent), Gest is gestational status (nullipara, primipara, multipara, multigravida), Age is maternal age in years, and $\epsilon$ is the error term. Results are displayed in Table 3.

```{r}
# Using recipes and workflows
# Define recipe for model
ML_recipe2 <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

@@

# Save regression output into table
saveRDS(table3, file = here::here("results", "tables", "table3_final.rds"))
```

@@

```{r}
# Using recipes and workflows
# Simple linear model
# Specify recipe
ML_recipe3 <- recipe(Weightgrams ~ Color, data = ML_linear) %>%
  step_relevel(Color, ref_level = "White")

@@

@@

@@

```{r}
# Using reipes and workflows
# Complex linear model
# Specify recipe
ML_recipe4 <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multipara") %>% # Set reference level for ModifiedStatus
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

@@

# Save regression output into table
saveRDS(table5, file = here("results", "tables", "table5_final.rds"))
```

>>>>

#rm(list=ls()) #clear the environment
```

@@

# Preparation for Model Assessment

First, I will create train and test data (75/25) to be able to evaluate the performance of our models. I will also set the seed to ensure reproducibility. For each model, I perform the following steps: 1) run the model on the train data; 2) perform cross-validation on the train data; 3) evaluate the model on the test data. The model outputs included in the final manuscript are on all the data.

```{r, message=FALSE,  warning=FALSE}
# Fix the random numbers by setting the seed 
# This enables the analysis to be reproducible when random numbers are used 
rngseed = 1234
set.seed(rngseed)

# Create training data
# Put 3/4 of the data into the training set 
data_split <- initial_split(ML_linear, prop = 3/4)

# Create data frames for the two sets:
train_data <- training(data_split)
test_data  <- testing(data_split)
```

# Model 1

@@

whereas BW is birthweight in grams, MA is maternal ancestry (Euro-descent, Afro-descent), and $\epsilon$ is the error term.

Full results are displayed in Table 2.

```{r}
# Model 1: Final model for paper using all data
# Model using recipes and workflows
# Define recipe for model
ML_recipe1 <- recipe(Weightgrams ~ ModifiedColor, data = ML_linear) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable

@@

# Save regression output into table
saveRDS(table2, file = here::here("results", "tables", "table2_final.rds"))
```
## Model 1 Performance

First we will evaluate the model on our training data.

```{r}
# Training model 1
# Model using recipes and workflows
# Define recipe for model
ML_recipe1_train <- recipe(Weightgrams ~ ModifiedColor, data = train_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable

# Specify model
ML_spec1_train <- linear_reg() %>% 
  set_engine("lm")

# Create work flow
ML_wf1_train <- workflow() %>%
  add_recipe(ML_recipe1_train) %>%
  add_model(ML_spec1_train)

# Fit the work flow
ML_fit1_train <- fit(ML_wf1_train, data = train_data)

# Summarize model fit with tidiers
ML_summary1_train <- summary(ML_fit1_train) # Summary of fit
tidy(ML_fit1_train, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit1_train) # Model-level statistics

# Produce predictions for R-squared
ML_fit1_pred_train <- predict(ML_fit1_train, new_data = train_data %>% select(-Weightgrams))
ML_fit1_pred_train

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit1_pred_train <- bind_cols(ML_fit1_pred_train, train_data %>% select(Weightgrams))
ML_fit1_pred_train

# Plot observed versus predicted
plot1_train <- ggplot(ML_fit1_pred_train, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot1_train

# Create metric set including RMSE and R-squared
metrics_fit1_train <- metric_set(rmse, rsq)
metrics_fit1_train(ML_fit1_pred_train, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit1_train <- as.data.frame(metrics_fit1_train(ML_fit1_pred_train, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit1_train

# Create a gtsummary table directly from the model
table2_train <-
  tbl_regression(ML_fit1_train, exponentiate = FALSE, intercept = TRUE, show_single_row = everything()) %>% 
  bold_p() %>% 
  add_glance_table(include = c(r.squared, adj.r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE)
table2_train

# Save regression output into table
saveRDS(table2_train, file = here::here("results", "tables", "table2_train.rds"))
```
Our results here are not that different from the model run on all the data. Our adjusted R-squared is 0.007, the same as the full model. The RMSE is 495, which is slightly lower than the full model (502), which shows better performance.

Now let's perform cross-validation on the training data.

```{r, message=FALSE,  warning=FALSE}
# CV model 1
# Set the seed for reproducibility
set.seed(rngseed)

# Create 10-fold cross-validation
folds <- vfold_cv(train_data, v = 10)
folds

# Define recipe for model
ML_recipe1_cv <- recipe(Weightgrams ~ ModifiedColor, data = train_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable

# Set model specification
ML_spec1_cv <- linear_reg() %>% set_engine("lm")

# Create and initialize workflow that bundles together model specification and formula.
ML_wf1_cv <- workflow() %>% 
  add_model(ML_spec1_cv) %>% 
  add_recipe(ML_recipe1_cv)
  fit_resamples(ML_wf1_cv, resamples = folds)

# Resample the model
ML_wf1_cv_rs <- fit_resamples(ML_wf1_cv, resamples = folds)
ML_wf1_cv_rs

# Get model metrics
ML_cv1_metrics <- collect_metrics(ML_wf1_cv_rs)

# Filter the data frame to include only rows with "rmse" and "rsq" in the ".metric" column
ML_cv1_metrics <- ML_cv1_metrics[ML_cv1_metrics$.metric %in% c("rmse", "rsq"), !(names(ML_cv1_metrics) %in% ".config")]
ML_cv1_metrics

# Save regression output into table
saveRDS(ML_cv1_metrics, file = here::here("results", "tables", "ML_cv1_metrics.rds"))
```
Performing cross-validation results in an RMSE that is the same as our training model (495), lower than our full model. The r-squared (0.02) is also higher than our training model (0.008) and our full model (0.007), demonstrating better model performance.

Finally, let's run the model on our test data.

```{r, message=FALSE,  warning=FALSE}
# Testing model 1
# Model using recipes and workflows
# Define recipe for model
ML_recipe1_test <- recipe(Weightgrams ~ ModifiedColor, data = test_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable

# Specify model
ML_spec1_test <- linear_reg() %>% 
  set_engine("lm")

# Create work flow
ML_wf1_test <- workflow() %>%
  add_recipe(ML_recipe1_test) %>%
  add_model(ML_spec1_test)

# Fit the work flow
ML_fit1_test <- fit(ML_wf1_test, data = test_data)

# Summarize model fit with tidiers
ML_summary1_test <- summary(ML_fit1_test) # Summary of fit
tidy(ML_fit1_test, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit1_test) # Model-level statistics

# Produce predictions for R-squared
ML_fit1_pred_test <- predict(ML_fit1_test, new_data = test_data %>% select(-Weightgrams))
ML_fit1_pred_test

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit1_pred_test <- bind_cols(ML_fit1_pred_test, test_data %>% select(Weightgrams))
ML_fit1_pred_test

# Plot observed versus predicted
plot1_test <- ggplot(ML_fit1_pred_test, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot1_test

# Create metric set including RMSE and R-squared
metrics_fit1_test <- metric_set(rmse, rsq)
metrics_fit1_test(ML_fit1_pred_test, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit1_test <- as.data.frame(metrics_fit1_test(ML_fit1_pred_test, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit1_test

# Create a gtsummary table directly from the model
table2_test <-
  tbl_regression(ML_fit1_test, exponentiate = FALSE, intercept = TRUE, show_single_row = everything()) %>% 
  bold_p() %>% 
  add_glance_table(include = c(r.squared, adj.r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE)
table2_test

# Save regression output into table
saveRDS(table2_test, file = here::here("results", "tables", "table2_test.rds"))
```

When running the model on our testing dataset, our results are not significant, given the small sample size.

# Model 2

@@

whereas BW is birthweight in grams, MA is maternal ancestry (Euro-descent, Afro-descent), Gest is gestational status (nulliparous or multiparous), Age is maternal age in years, and $\epsilon$ is the error term. Results are displayed in Table 3.

```{r}
# Final model 2 for paper using all data
# Using recipes and workflows
# Define recipe for model
ML_recipe2 <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

# Produce predictions for R-squared
ML_fit2_pred <- predict(ML_fit2, new_data = ML_linear %>% select(-Weightgrams))
ML_fit2_pred

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit2_pred <- bind_cols(ML_fit2_pred, ML_linear %>% select(Weightgrams))
ML_fit2_pred

# Plot observed versus predicted
plot2 <- ggplot(ML_fit2_pred, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot2

# Create metric set including RMSE and R-squared
metrics_fit2 <- metric_set(rmse, rsq)
metrics_fit2(ML_fit2_pred, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit2 <- as.data.frame(metrics_fit2(ML_fit2_pred, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit2

@@

# Save regression output into table
saveRDS(table3, file = here::here("results", "tables", "table3_final.rds"))
```
Our RMSE is 498 and our r-squared is 0.02.

## Model 2 Performance

Now let's run the model on our training data.

```{r}
# Training model 2
# Using recipes and workflows
# Define recipe for model
ML_recipe2_train <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = train_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Define model specification
ML_spec2_train <- linear_reg() %>% 
  set_engine("lm") %>%
  set_mode("regression")

# Create workflow
ML_wf2_train <- workflow() %>%
  add_recipe(ML_recipe2_train) %>%
  add_model(ML_spec2_train)

# Fit the workflow
ML_fit2_train <- fit(ML_wf2_train, data = train_data)

# Summarize model fit with tidiers
ML_lm2_summary_train <- summary(ML_fit2_train) # Summary of fit
tidy(ML_fit2_train, conf.int = TRUE) %>% mutate(p.value = round(p.value, 5)) # Coefficients and confidence intervals with p-values rounded to 5 decimal places
glance(ML_fit2_train) # Model-level statistics

# Produce predictions for R-squared
ML_fit2_pred_train <- predict(ML_fit2_train, new_data = train_data %>% select(-Weightgrams))
ML_fit2_pred_train

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit2_pred_train <- bind_cols(ML_fit2_pred_train, train_data %>% select(Weightgrams))
ML_fit2_pred_train

# Plot observed versus predicted
plot2_train <- ggplot(ML_fit2_pred_train, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot2_train

# Create metric set including RMSE and R-squared
metrics_fit2_train <- metric_set(rmse, rsq)
metrics_fit2_train(ML_fit2_pred_train, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit2_train <- as.data.frame(metrics_fit2_train(ML_fit2_pred_train, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit2_train

# Create a gtsummary table directly from the model
table3_train <- pull_workflow_fit(ML_fit2_train) %>%
  tbl_regression(exponentiate = FALSE, intercept = TRUE) %>%
  add_glance_table(include = c(r.squared, adj.r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE) %>%
  bold_p()
table3_train

# Save regression output into table
saveRDS(table3_train, file = here::here("results", "tables", "table3_train.rds"))
```
Our RMSE is 490 and our r-squared is 0.03, a slightly better performance than the full model.

Now let's run cross-validation.

```{r, message=FALSE,  warning=FALSE}
# CV model 2
# Set the seed for reproducibility
set.seed(rngseed)

# Create 10-fold cross-validation
folds <- vfold_cv(train_data, v = 10)
folds

# Define recipe for model
ML_recipe2_cv <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = train_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Set model specification
ML_spec2_cv <- linear_reg() %>% set_engine("lm")

# Create and initialize workflow that bundles together model specification and formula.
ML_wf2_cv <- workflow() %>% 
  add_model(ML_spec2_cv) %>% 
  add_recipe(ML_recipe2_cv)
  fit_resamples(ML_wf2_cv, resamples = folds)

# Resample the model
ML_wf2_cv_rs <- fit_resamples(ML_wf2_cv, resamples = folds)
ML_wf2_cv_rs

# Get model metrics
ML_cv2_metrics <- collect_metrics(ML_wf2_cv_rs)

# Filter the data frame to include only rows with "rmse" and "rsq" in the ".metric" column
ML_cv2_metrics <- ML_cv2_metrics[ML_cv2_metrics$.metric %in% c("rmse", "rsq"), !(names(ML_cv2_metrics) %in% ".config")]
ML_cv2_metrics

# Save regression output into table
saveRDS(ML_cv2_metrics, file = here::here("results", "tables", "ML_cv2_metrics.rds"))
```
Here, our RMSE is 458, an improvement from both the full model and the training model, and our r-squared is 0.03, the same as our training model.

Finally, let's run our model on the test data.

```{r}
# Testing model 2
# Using recipes and workflows
# Define recipe for model
ML_recipe2_test <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = test_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Define model specification
ML_spec2_test <- linear_reg() %>% 
  set_engine("lm") %>%
  set_mode("regression")

# Create workflow
ML_wf2_test <- workflow() %>%
  add_recipe(ML_recipe2_test) %>%
  add_model(ML_spec2_test)

# Fit the workflow
ML_fit2_test <- fit(ML_wf2_test, data = test_data)

# Summarize model fit with tidiers
ML_lm2_summary_test <- summary(ML_fit2_test) # Summary of fit
tidy(ML_fit2_test, conf.int = TRUE) %>% mutate(p.value = round(p.value, 5)) # Coefficients and confidence intervals with p-values rounded to 5 decimal places
glance(ML_fit2_test) # Model-level statistics

# Produce predictions for R-squared
ML_fit2_pred_test <- predict(ML_fit2_test, new_data = test_data %>% select(-Weightgrams))
ML_fit2_pred_test

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit2_pred_test <- bind_cols(ML_fit2_pred_test, test_data %>% select(Weightgrams))
ML_fit2_pred_test

# Plot observed versus predicted
plot2_test <- ggplot(ML_fit2_pred_test, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot2_test

# Create metric set including RMSE and R-squared
metrics_fit2_test <- metric_set(rmse, rsq)
metrics_fit2_test(ML_fit2_pred_test, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit2_test <- as.data.frame(metrics_fit2_test(ML_fit2_pred_test, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit2_test

# Create a gtsummary table directly from the model
table3_test <- pull_workflow_fit(ML_fit2_test) %>%
  tbl_regression(exponentiate = FALSE, intercept = TRUE) %>%
  add_glance_table(include = c(r.squared, adj.r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE) %>%
  bold_p()
table3_test

# Save regression output into table
saveRDS(table3_test, file = here::here("results", "tables", "table3_test.rds"))
```
Running the model on our test data, we see that the only significant result is for the nulliparous status. All other results are insignificant.

# Model 3

@@

```{r}
# Final model 3 for paper using all data
# Using recipes and workflows
# Simple linear model
# Specify recipe
ML_recipe3 <- recipe(Weightgrams ~ Color, data = ML_linear) %>%
  step_relevel(Color, ref_level = "White")

@@

# Produce predictions for R-squared
ML_fit3_pred <- predict(ML_fit3, new_data = ML_linear %>% select(-Weightgrams))
ML_fit3_pred

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit3_pred <- bind_cols(ML_fit3_pred, ML_linear %>% select(Weightgrams))
ML_fit3_pred

# Plot observed versus predicted
plot3 <- ggplot(ML_fit3_pred, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot3

# Create metric set including RMSE and R-squared
metrics_fit3 <- metric_set(rmse, rsq)
metrics_fit3(ML_fit3_pred, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit3 <- as.data.frame(metrics_fit3(ML_fit3_pred, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit3

@@

Our RMSE is 502 and our r-squared is 0.007.

## Model 3 Performance

Let's now evaluate the model on our training data.

```{r, message=FALSE,  warning=FALSE}
# Training model 3
# Using recipes and workflows
# Specify recipe
ML_recipe3_train <- recipe(Weightgrams ~ Color, data = train_data) %>%
  step_relevel(Color, ref_level = "White")

# Specify model
ML_spec3_train <- linear_reg() %>%
  set_engine("lm")

# Create workflow
ML_wf3_train <- workflow() %>%
  add_recipe(ML_recipe3_train) %>%
  add_model(ML_spec3_train)

# Fit the workflow
ML_fit3_train <- fit(ML_wf3_train, data = train_data)

# Summarize model fit with tidiers and create a gtsummary table directly from the model
ML_summary3_train <- summary(ML_fit3_train)
tidy(ML_fit3_train, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit3_train) # Model-level statistics

# Produce predictions for R-squared
ML_fit3_pred_train <- predict(ML_fit3_train, new_data = train_data %>% select(-Weightgrams))
ML_fit3_pred_train

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit3_pred_train <- bind_cols(ML_fit3_pred_train, train_data %>% select(Weightgrams))
ML_fit3_pred_train

# Plot observed versus predicted
plot3_train <- ggplot(ML_fit3_pred_train, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot3_train

# Create metric set including RMSE and R-squared
metrics_fit3_train <- metric_set(rmse, rsq)
metrics_fit3_train(ML_fit3_pred_train, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit3_train <- as.data.frame(metrics_fit3_train(ML_fit3_pred_train, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit3_train

# Create a gtsummary table from model
table4_train <- tbl_regression(ML_fit3_train, exponentiate = FALSE, intercept = TRUE) %>%
  bold_p() %>%
  add_glance_table(include = c(r.squared, adj.r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE)
table4_train

# Save regression output into table
saveRDS(table4_train, file = here("results", "tables", "table4_train.rds"))
```

Our RMSE is 495 and our r-squared is 0.009, slightly better performance than the full model. Our results are not significantly different from the full model.

Now let's run cross-validation.

```{r, message=FALSE,  warning=FALSE}
# CV model 3
# Set the seed for reproducibility
set.seed(rngseed)

# Create 10-fold cross-validation
folds <- vfold_cv(train_data, v = 10)
folds

# Define recipe for model
ML_recipe3_cv <- recipe(Weightgrams ~ Color, data = train_data) %>%
  step_relevel(Color, ref_level = "White")

# Set model specification
ML_spec3_cv <- linear_reg() %>% set_engine("lm")

# Create and initialize workflow that bundles together model specification and formula.
ML_wf3_cv <- workflow() %>% 
  add_model(ML_spec3_cv) %>% 
  add_recipe(ML_recipe3_cv)
  fit_resamples(ML_wf3_cv, resamples = folds)

# Resample the model
ML_wf3_cv_rs <- fit_resamples(ML_wf3_cv, resamples = folds)
ML_wf3_cv_rs

# Get model metrics
collect_metrics(ML_wf3_cv_rs)

# Get model metrics
ML_cv3_metrics <- collect_metrics(ML_wf3_cv_rs)

# Filter the data frame to include only rows with "rmse" and "rsq" in the ".metric" column
ML_cv3_metrics <- ML_cv3_metrics[ML_cv3_metrics$.metric %in% c("rmse", "rsq"), !(names(ML_cv3_metrics) %in% ".config")]
ML_cv3_metrics

# Save regression output into table
saveRDS(ML_cv3_metrics, file = here::here("results", "tables", "ML_cv3_metrics.rds"))
```
Using cross-validation, our RMSE is 495 and our r-squared is 0.02, a similar performance to our training model.

Finally, let's run our model on the test data.

```{r, message=FALSE, warning=FALSE}
# Testing model 3
# Using recipes and workflows
# Specify recipe
ML_recipe3_test <- recipe(Weightgrams ~ Color, data = test_data) %>%
  step_relevel(Color, ref_level = "White")

# Specify model
ML_spec3_test <- linear_reg() %>%
  set_engine("lm")

# Create workflow
ML_wf3_test <- workflow() %>%
  add_recipe(ML_recipe3_test) %>%
  add_model(ML_spec3_test)

# Fit the workflow
ML_fit3_test <- fit(ML_wf3_test, data = test_data)

# Summarize model fit with tidiers and create a gtsummary table directly from the model
ML_summary3_test <- summary(ML_fit3_test)
tidy(ML_fit3_test, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit3_test) # Model-level statistics

# Produce predictions for R-squared
ML_fit3_pred_test <- predict(ML_fit3_test, new_data = test_data %>% select(-Weightgrams))
ML_fit3_pred_test

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit3_pred_test <- bind_cols(ML_fit3_pred_test, test_data %>% select(Weightgrams))
ML_fit3_pred_test

# Plot observed versus predicted
plot3_test <- ggplot(ML_fit3_pred_test, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot3_test

# Create metric set including RMSE and R-squared
metrics_fit3_test <- metric_set(rmse, rsq)
metrics_fit3_test(ML_fit3_pred_test, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit3_test <- as.data.frame(metrics_fit3_test(ML_fit3_pred_test, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit3_test

# Create a gtsummary table from model
table4_test <- tbl_regression(ML_fit3_test, exponentiate = FALSE, intercept = TRUE) %>%
  bold_p() %>%
  add_glance_table(include = c(r.squared, adj.r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE)
table4_test

# Save regression output into table
saveRDS(table4_test, file = here("results", "tables", "table4_test.rds"))
```

Much like in the other testing models, our results are not significant (except the intercept). Our performance is also worse (RMSE of 518 and r-squared of 0.005).

# Model 4

@@

```{r}
# Final model 4 for paper using all data
# Using recipes and workflows
# Complex linear model
# Specify recipe
ML_recipe4 <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") %>% # Set reference level for ModifiedStatus
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

# Produce predictions for R-squared
ML_fit4_pred <- predict(ML_fit4, new_data = ML_linear %>% select(-Weightgrams))
ML_fit4_pred

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit4_pred <- bind_cols(ML_fit4_pred, ML_linear %>% select(Weightgrams))
ML_fit4_pred

# Plot observed versus predicted
plot4 <- ggplot(ML_fit4_pred, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot4

# Create metric set including RMSE and R-squared
metrics_fit4 <- metric_set(rmse, rsq)
metrics_fit4(ML_fit4_pred, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit4 <- as.data.frame(metrics_fit4(ML_fit4_pred, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit4

@@

# Save regression output into table
saveRDS(table5, file = here("results", "tables", "table5_final.rds"))
```
Our RMSE is 498 and our r-squared is 0.02.

## Model 4 Performance

Let's now evaluate the model on our training data.

```{r, message=FALSE,  warning=FALSE}
# Model 4 training
# Using recipes and workflows
# Complex linear model
# Specify recipe
ML_recipe4_train <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = train_data) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") %>% # Set reference level for ModifiedStatus
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Specify model
ML_spec4_train <- linear_reg() %>%
  set_engine("lm")

# Create workflow
ML_wf4_train <- workflow() %>%
  add_recipe(ML_recipe4_train) %>%
  add_model(ML_spec4_train)

# Fit the workflow
ML_fit4_train <- fit(ML_wf4_train, data = train_data)

# Summarize model fit with tidiers and create a gtsummary table directly from the model
ML_summary4_train <- summary(ML_fit4_train)
tidy(ML_fit4_train, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit4_train) # Model-level statistics

# Produce predictions for R-squared
ML_fit4_pred_train <- predict(ML_fit4_train, new_data = train_data %>% select(-Weightgrams))
ML_fit4_pred_train

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit4_pred_train <- bind_cols(ML_fit4_pred_train, train_data %>% select(Weightgrams))
ML_fit4_pred_train

# Plot observed versus predicted
plot4_train <- ggplot(ML_fit4_pred_train, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot4_train

# Create metric set including RMSE and R-squared
metrics_fit4_train <- metric_set(rmse, rsq)
metrics_fit4_train(ML_fit4_pred_train, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit4_train <- as.data.frame(metrics_fit4_train(ML_fit4_pred_train, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit4_train
